---
# Copyright 2025 Snowflake Inc. 
# SPDX-License-Identifier: Apache-2.0
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.
title: "IntroAnalysis"
output:
  pdf_document:
    toc: true
    number_sections: true
  html_document:
    toc: true
    number_sections: true
date: "2025-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r libraries, echo=FALSE, results=FALSE, warning=FALSE, message=FALSE}
defaultW <- getOption("warn") # suppress all informational messages during pkg load
options(warn=-1)
library(ggplot2)
library(tidyr)
library(dplyr)
library(arrow)
options(warn=defaultW)
```

# Dependencies

The R code in this repository is provided in [R Markdown](https://rmarkdown.rstudio.com/) files. 
This is a notebook format that is well supported by IDEs such as [R Studio](https://www.rstudio.com).

## Packages

The animation depends on ffmpeg to stitch together frames into an animation.  If you are installing R packages from source, you may also need to install a few packages.
On MacOS, you should install:

```
brew install ffmpeg
brew install gdal
brew install units
```

## Libraries

The following dependent libraries must be installed in your version of R to run the notebooks here.

```{r libdependencies, eval=FALSE}
install.packages(c("tidyr", "dplyr", "ggplot2", "ggpattern", "arrow", "zoo",
                   "lubridate", "cowplot"))
```

# Reading the Data

## R

This provided parquet file can be read into R or Python with the relevant parquet libraries, or queried directly with Snowflake, DuckDB, or other tools.

We start by examing the first few rows of the data.  The total dataset has 524,832 rows.

```{r readdata}
library(arrow)
data <- read_parquet("./hourly_normalized.parquet")
head(data)
dim(data)
```
## Python

See [IntroAnalysis-py.ipynb](./IntroAnalysis-py.ipynb) for a Jupyter notebook to read in and plot this dataset with Python.

## DuckDB

```
duckdb
D select count(*) from read_parquet('./hourly_normalized.parquet');
524832
```

# Visualization

## Complete Dataset

The following graph shows every data point in the dataset.  There is a separate panel for each region, and then a separate colored line for each instance type within the region, showing the relative demand for different VM types across time.


```{r visualizeall}
ggplot(data, aes(x=USAGE_HOUR, y=NORM_USAGE, color=INSTANCE_TYPE)) +
  geom_line() + theme_bw() +
  theme(axis.text=element_text(size=15), axis.title=element_text(size=15),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15),
        strip.text=element_text(size=15)) +
  ggtitle("Normalized VM Demand from 12 Instance Types Across 4 Regions") +
  facet_wrap(. ~ REGION_NUM)
```


```{r visualizeallpng, echo=FALSE}
png("timeseries.png", width=900, height=600)
ggplot(data, aes(x=USAGE_HOUR, y=NORM_USAGE, color=INSTANCE_TYPE)) +
  geom_line() + theme_bw() +
  theme(axis.text=element_text(size=15), axis.title=element_text(size=15),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15),
        strip.text=element_text(size=15)) +
  ggtitle("Normalized VM Demand from 12 Instance Types Across 4 Regions") +
  facet_wrap(. ~ REGION_NUM)
dev.off()
```


## Subset of 2 specific instance types

```{r visualizesubsets}
data.sub <- data %>% subset(INSTANCE_TYPE %in% c("A", "B"))
ggplot(data.sub, aes(x=USAGE_HOUR, y=NORM_USAGE, color=INSTANCE_TYPE)) +
  geom_line() + theme_bw() +
  theme(axis.text=element_text(size=15), axis.title=element_text(size=15),
        legend.text = element_text(size=15), legend.title = element_blank(),
        legend.position="bottom", strip.text=element_text(size=15)) +
  facet_wrap(. ~ REGION_NUM)
```

## Subsetting to a specific region and then aggregating all VM types together for a regional demand timeseries

```{r visualizeaggregate}
data.2 <- data %>% subset(REGION_NUM==2) %>% group_by(USAGE_HOUR) %>%
  summarise(across(NORM_USAGE, sum))
head(data.2)
ggplot(data.2, aes(x=USAGE_HOUR, y=NORM_USAGE)) + geom_line()
```

## Subset of a week to examine weekly pattern

```{r weeklypattern}
data.sub <- data.2 %>%
  subset(USAGE_HOUR >= as.POSIXct("2024-01-07", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2024-01-14", tz="UTC"))
ggplot(data.sub, aes(x=USAGE_HOUR, y=NORM_USAGE)) + geom_line()
data.sub
```

