---
# Copyright 2025 Snowflake Inc. 
# SPDX-License-Identifier: Apache-2.0
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.
title: "Figures: Optimal Commitment Levels"
author: "Murray Stokely"
email: "murray.stokely@snowflake.com"
output:
  pdf_document:
    toc: true
    number_sections: true
  html_document:
    toc: true
    number_sections: true
date: "2024-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, results=FALSE, warning=FALSE, message=FALSE}
defaultW <- getOption("warn") # suppress all informational messages during pkg load
options(warn=-1)
library(zoo)
library(ggplot2)
library(tidyr)
library(dplyr)
library(arrow)
library(cowplot)
library(scales)
library(ggpattern) # for thatched area plots to not rely on color printing
options(warn=defaultW)
```

# Introduction

This file includes figures and analysis for Section 3.2 on Setting Optimal
Commitment Level for Periodic Demand.

We start by reading in the dataset.

```{r}
data <- read_parquet("../hourly_normalized.parquet")
```

We then aggregating the different regions and SKUs together to a single timeseries of total VM demand and then normalize this demand to a 100 unit peak over the period.

```{r}
data.all <- data %>% group_by(USAGE_HOUR) %>% summarise(across(NORM_USAGE, sum))
data.all$NORM_USAGE = 100 * data.all$NORM_USAGE / max(data.all$NORM_USAGE)
```

We then create a 2-week subset of the data from January 2024, and check that we have the data we expect.

```{r}
data.sub <- data.all %>% 
  subset(USAGE_HOUR >= as.POSIXct("2024-01-07", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2024-01-21", tz="UTC"))
data.sub$NORM_USAGE = 100 * data.sub$NORM_USAGE / max(data.sub$NORM_USAGE)
dim(data.sub)
head(data.sub)
range(data.sub$USAGE_HOUR)
ggplot(data.sub, aes(x=USAGE_HOUR, y=NORM_USAGE)) + geom_line()
```

# Visualization

In order to illustrate the impact of savings plans on the amount of VM demand
that is (1) covered by a savings plan, (2) purchased with on-demand rates, and (3) wasted as an unused savings plan,
we introduce a simple 3-color area visualization with time on the x-axis and normalized cost on the y-axis.

Note that since the y-axis is cost, instead of VM instance hours, the more of the demand that is covered at expensive on-demand rates means the higher the y-axis will be.

```{r generateboxes}
GenerateBoxes <- function(df, on.demand.premium=2.1) {
  # GenerateBoxes - Create data.frame with pricing
  # Args:
  #   df: A data.frame with columns
  #     USAGE_HOUR: POSIXct hourly time
  #     NORM_USAGE: The VM demand at that hour
  # 3 boxes each time range - ununused, used with sp/ri coverage,
  #   used above sp/ri coverage level.
  data.boxes <- data.frame(
    xmin = rep(head(df$USAGE_HOUR, -1), 3),
    xmax = rep(tail(df$USAGE_HOUR, -1), 3),
    Pricing = c(rep("SP Commitment", length(head(df$USAGE_HOUR, -1))),
                rep("Unused Commitment", length(head(df$USAGE_HOUR, -1))),
                rep("On Demand", length(head(df$USAGE_HOUR, -1)))),
    ymin = c(rep(0, length(head(df$USAGE_HOUR, -1))),
             head(ifelse(df$NORM_USAGE < df$SPRI_LEVEL, df$NORM_USAGE,
                         df$SPRI_LEVEL), -1),
             head(df$SPRI_LEVEL, -1)),
    ymax = c(head(ifelse(df$NORM_USAGE < df$SPRI_LEVEL, df$NORM_USAGE,
                         df$SPRI_LEVEL), -1),
             head(df$SPRI_LEVEL,-1),
             head(ifelse(df$NORM_USAGE > df$SPRI_LEVEL,
                         df$SPRI_LEVEL +
                           (df$NORM_USAGE - df$SPRI_LEVEL)*on.demand.premium,
                         df$SPRI_LEVEL), -1)))
  
  return(data.boxes)
}
```

```{r plotboxes}
AnnotateSPRILevel <- function(df, sp.level) {
  df$SPRI_LEVEL = sp.level
  return(df)
}

PlotBoxes <- function(df, sp.level=NULL, title=FALSE, ylim=NULL) {
  days <- c("S", "M", "T", "W", "R", "F", "S")
  p <- ggplot(df) +
           geom_rect_pattern(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax,
                                 fill=Pricing, pattern=Pricing),
                             colour=NA, pattern_size=0.25,
                             pattern_spacing=0.02) +
           theme_bw() +
           theme(axis.text=element_text(size=15),
                 axis.title=element_text(size=15),
                 legend.text = element_text(size=15),
                 legend.title = element_blank(),
                 legend.position="bottom",
                 strip.text=element_text(size=15)) +
           ylab("Cost") + xlab("") +
           scale_pattern_manual(values=c("none", "none", "stripe")) +
           scale_x_continuous(breaks=seq(from=as.POSIXct("2024-06-02 12:00:00"),
                                         to=as.POSIXct("2024-06-15 12:00:00"),
                                         length.out=14),
                              labels=rep(days,2)) +
           scale_fill_manual(values=c("orange", "darkgreen", "red"))
  if (title) {
    if (!is.null(sp.level)) {
      cost.premium  <- TotalCostPremium(df)
      p <- p + ggtitle(paste0("c=", round(sp.level, 1), " ", "C(c)=", round(cost.premium, 0)))
    }
  }
  if (!is.null(ylim)) {
    p <- p+ylim(ylim)
  }
  return(p)
}
```
  

```{r}
# Set the Savings Plan commitment level to halfway between the max and min
# demand over the time period.
data.sub <- AnnotateSPRILevel(data.sub,
                              (max(data.sub$NORM_USAGE) +
                                 min(data.sub$NORM_USAGE))/2)
df <- GenerateBoxes(data.sub)
head(df)
PlotBoxes(df)
```

# Optimization

The final step is to iterate over a number of possible SP commitment levels
to compute the minimum cost option given the VM demand curve.

```{r}
TotalCostPremium <- function(df) {
  # TotalCost - Computes the Total Cost Premium
  #
  # Args:
  #   df - A data.frame with 3 areas - unused, covered, on-demand.
  # Returns
  #   Cost premium

  l1 <- subset(df, Pricing == "SP Commitment")
  l2 <- subset(df, Pricing == "Unused Commitment")
  l3 <- subset(df, Pricing == "On Demand")
  plan.cost <- sum(l1$ymax - l1$ymin)
  unused.cost <- sum(l2$ymax - l2$ymin)
  ondemand.cost <- sum(l3$ymax - l3$ymin)
  total.cost <- plan.cost + unused.cost + ondemand.cost
  return(total.cost)
}


# findOptimalSPLevel - Iteratively identify lowest cost Savings Plan level
# Args:
#   df: a data.frame containing 3 columns
#.     USAGE_HOUR - timestamp
#.     NORM_USAGE - normalized usage
#.  steps: The number of steps to iterate through the possible SP levels
# Returns:
#   A number corresponding to the optimal SP level to minimize cost.
findOptimalSPLevel <- function(df, steps=9) {
  min.demand <- min(df$NORM_USAGE)
  max.demand <- max(df$NORM_USAGE)
  total.days <- as.numeric(max(df$USAGE_HOUR) - min(df$USAGE_HOUR))

  fulldf <- NULL
  plts <- list()
  i<-1
  for (sp.level in seq(min.demand, max.demand, length=steps)) {
    df.ann <- AnnotateSPRILevel(df, sp.level)
    df.boxes <- GenerateBoxes(df.ann)
    df.boxes$sp.level.label <- sp.level
    cost.premium  <- TotalCostPremium(df.boxes)
    df.boxes$cost.premium <- cost.premium
    df.boxes$panel.title <- paste0(i, ". c=", round(sp.level, 1), " ",
                                   "C(c)=", round(cost.premium, 0))
    i <- i + 1

    if (is.null(fulldf)) {
      fulldf <- df.boxes
    } else {
      fulldf <- rbind(fulldf, df.boxes)
    }
  }
  return(fulldf)
}

findMinSPLevel <- function(df, steps=9) {
  min.demand <- min(df$NORM_USAGE)
  max.demand <- max(df$NORM_USAGE)
  total.days <- as.numeric(max(df$USAGE_HOUR) - min(df$USAGE_HOUR))

  lowest.cost <- NULL
  lowest.sp <- NULL
  i<-1
  for (sp.level in seq(min.demand, max.demand, length=steps)) {
    df.ann <- AnnotateSPRILevel(df, sp.level)
    df.boxes <- GenerateBoxes(df.ann)
    df.boxes$sp.level.label <- sp.level
    cost.premium  <- TotalCostPremium(df.boxes)
    if (is.null(lowest.cost)) {
      lowest.cost <- cost.premium
      lowest.sp <- sp.level
    } else if (cost.premium < lowest.cost) {
      lowest.cost <- cost.premium
      lowest.sp <- sp.level
    }
  }
  return(lowest.sp)
}

FindCost <- function(df, sp.level) {
  df.ann <- AnnotateSPRILevel(df, sp.level)
  df.boxes <- GenerateBoxes(df.ann)
  df.boxes$sp.level.label <- sp.level
  return(TotalCostPremium(df.boxes))
}

fulldf.1 <- findOptimalSPLevel(data.sub)
lowest.1 <- findMinSPLevel(data.sub, 100)
cost.1 <- FindCost(data.sub, lowest.1)
print(lowest.1)
print(cost.1)
```

We use a 3x3 matrix of visualizations with different savings plan commitment levels to illustrate how the minimum cost is found.

```{r plot3by3}
Plot3x3 <- function(df, start.date="2024-01-07", end.date="2024-01-21", weekdays=FALSE) {
  days <- c("S", "M", "T", "W", "R", "F", "S")
  p <- ggplot(df) +
    geom_rect_pattern(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax,
                          fill=Pricing, pattern=Pricing),
                      colour=NA, pattern_size=0.25, pattern_spacing=0.02) +
    theme_bw() +
    theme(axis.text=element_text(size=15), axis.title=element_text(size=15),
          legend.text = element_text(size=15), legend.title = element_blank(),
          legend.position="bottom", strip.text=element_text(size=15)) +
    ylab("Cost") + xlab("") +
    scale_pattern_manual(values=c("none", "none", "stripe"))
  if (weekdays) {
    p <- p + scale_x_continuous(breaks=seq(from=as.POSIXct(start.date, tz="UTC"),
                                  to=as.POSIXct(end.date, tz="UTC"),
                                  length.out=14),
                      labels=rep(days,2))
  } else {
    #p <- p + scale_x_datetime(breaks = "1 week",
    #     minor_breaks = "1 day", labels=date_format("%b-%d"))
        p <- p + scale_x_continuous(breaks=c(
                                          as.POSIXct("2023-12-17", tz="UTC"),
                                          as.POSIXct("2023-12-24", tz="UTC")),
                                 labels=date_format("%b-%d"))
  }
  p <- p +
    scale_fill_manual(values=c("orange", "darkgreen", "red")) +
    facet_wrap(~panel.title)
  return(p)
}
```

Then we generate output PDF and PNGs:

## Figure 4

```{r}
Plot3x3(fulldf.1, weekdays=TRUE)
```

```{r echo=FALSE, eval=FALSE}
png("3x3.png", width=900, height=750)
Plot3x3(fulldf.1)
dev.off()
```

```{r echo=FALSE, eval=FALSE}
pdf("3x3.pdf", width=9, height=7.5)
Plot3x3(fulldf.1)
dev.off()
```

# Comparison of Different Commitment Levels

Lets compare what happens if we use 1-week into the future or 2 weeks, for example, in the week preceding a major seasonal effect.

We start by subsetting the data to look only at the last two weeks of 2023 including the holiday slump and then normalize demand to a maximum of 100 units during this time.

```{r}
data.tmp <- data.all %>% 
  subset(USAGE_HOUR >= as.POSIXct("2023-12-17", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2023-12-31", tz="UTC"))
# Re-normalize just these 4 weeks to 100
data.tmp$NORM_USAGE <- data.tmp$NORM_USAGE * (100/max(data.tmp$NORM_USAGE))
```

## Figure 8.

We take a further subset of only the first week from 12/17 to 12/24, and find the minimum SP level for that first week, as well as the minimum SP level for both weeks.

```{r}
data.tmp.1 <- data.tmp %>% subset(USAGE_HOUR < as.POSIXct("2023-12-24", tz="UTC"))
o.1 <-findMinSPLevel(data.tmp.1, steps=100)
o.2 <- findMinSPLevel(data.tmp, steps=100)
```


```{r}
data.fig8.2 <- AnnotateSPRILevel(data.tmp, o.2)

data.fig8.1 <- AnnotateSPRILevel(data.tmp, o.1)
df.1 <- GenerateBoxes(data.fig8.1)
PlotBoxes(df.1, sp.level=o.1, title=TRUE)

df.2 <- GenerateBoxes(data.fig8.2)
PlotBoxes(df.2, sp.level=o.2, title=TRUE) #, ylim=c(0,4600))
```
## Figure 8

```{r}
df.1$panel.title <- paste0("(a) c=", round(o.1, 1), " C(c)=", round(TotalCostPremium(df.1), 0))
df.2$panel.title <- paste0("(b) c=", round(o.2, 1), " C(c)=", round(TotalCostPremium(df.2), 0))
Plot3x3(rbind(df.1, df.2), start.date="2023-12-17", end.date="2023-12-31")
```


```{r echo=FALSE, eval=FALSE}
pdf("forecastw12.pdf", width=9, height=3.5)
Plot3x3(rbind(df.21, df.2), start.date="2023-12-17", end.date="2023-12-31")
dev.off()
```


```{r echo=FALSE, eval=FALSE}
png("forecastw12.png", width=900, height=350)
Plot3x3(rbind(df.21, df.2), start.date="2023-12-17", end.date="2023-12-31")
dev.off()
```
