---
# Copyright 2025 Snowflake Inc. 
# SPDX-License-Identifier: Apache-2.0
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.
title: "Figures: Snowflake Workload Timeseries Patterns"
author: "Murray Stokely"
email: "murray.stokely@snowflake.com"
output:
  pdf_document:
    toc: true
    number_sections: true
  html_document:
    toc: true
    number_sections: true
date: "2024-10-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, results=FALSE, warning=FALSE, message=FALSE}
defaultW <- getOption("warn")
options(warn=-1)
library(zoo)
library(ggplot2)
library(tidyr)
library(dplyr)
library(arrow)
library(cowplot)
library(lubridate)
options(warn=defaultW)
```

# Introduction

This file includes figures and analysis for Section 2.2 on the Snowflake Workload User Demand Patterns.  All figures and analysis are generated from the public dataset.

# Full-Granularity Aggregated Hourly Timeseries

All regions and SKU types aggregated together at full hourly granularity and plotted over the full time range of the data set.

```{r}
data <- read_parquet("../hourly_normalized.parquet")
data.all <- data %>% group_by(USAGE_HOUR) %>% summarise(across(NORM_USAGE, sum))
head(data.all)
ggplot(data.all, aes(x=USAGE_HOUR, y=NORM_USAGE)) + geom_line()
```

## Normalized Full-Granularity Aggregated Hourly Timeseries

All regions and SKU types aggregated together at full hourly granularity and plotted over the full time range of the data set.  Normalized with the largest data point at 100.

```{r}
data.p1 <- data %>% group_by(USAGE_HOUR) %>%
  summarise(across(NORM_USAGE, sum))
data.p1$TOTAL_COMPUTE = data.p1$NORM_USAGE
data.p1$day <- as.Date(data.p1$USAGE_HOUR)
data.p1$TOTAL_COMPUTE = (100*data.p1$TOTAL_COMPUTE) / max(data.p1$TOTAL_COMPUTE)
p1 <- ggplot(data.p1, aes(x=USAGE_HOUR, y=TOTAL_COMPUTE)) +
  geom_line(color="#5471AB") + theme_bw() +
  theme(axis.title.x=element_text(size=16), axis.title.y=element_text(size=16),
        axis.text=element_text(size=14)) +
  xlab("Year") + ylab("Normalized\nVM Instance Demand")
old.mai <- par("mai")
par("mai"= c(1,0.9,2,0.4))
p1
```

## Normalized Timeseries of the Daily Means

Plot of the normalized daily means.

```{r}
data.p2 <- data.p1 %>% group_by(day) %>%
  summarize(DAILY_MEAN = mean(TOTAL_COMPUTE))
data.p2$DAILY_MEAN = (100*data.p2$DAILY_MEAN) / max(data.p2$DAILY_MEAN)
p2 <- ggplot(data.p2, aes(x=day, y=DAILY_MEAN)) + geom_line(color="#5471AB") +
  theme_bw() + theme(axis.title.x=element_text(size=16),
                     axis.title.y=element_text(size=16),
                     axis.text=element_text(size=14)) +
  xlab("Year") + ylab("Normalized\nVM Instance Demand")
old.mai <- par("mai")
par("mai"= c(1,0.9,2,0.4))
p2
```

# Weekly Pattern Analysis

We trim the start and end of our timeseries toalign our dataset with Sunday through Saturday weeks so we can split it up in 7 day chunks and look at the distribution of weekly patterns.

We then generate timeseries of the maximum, minimum, and mean of each day.

```{r}
all <- data.p1 %>% subset(USAGE_HOUR >= as.POSIXct("2021-02-07", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2024-01-28", tz="UTC"))
range(all$USAGE_HOUR)
all.ts <- zoo(all$TOTAL_COMPUTE, all$USAGE_HOUR)
# Compute the maximumes over each day.
all.maxdaily <- rollapply(all.ts, width=24, by=24, FUN=max)
all.meandaily <- rollapply(all.ts, width=24, by=24, FUN=mean)
all.mindaily <- rollapply(all.ts, width=24, by=24, FUN=min)
plot(all.ts)
plot(all.maxdaily)
all.meandaily <- 100*all.meandaily/(max(all.meandaily))
plot(all.meandaily)
```

Next we plot the first two weeks of our dataset to examine the different
weekly periodicity.

```{r weeklysubsets}
range(all$USAGE_HOUR)
head(all)
plot(all$TOTAL_COMPUTE[1:(24*7)], type="l")
plot(all$TOTAL_COMPUTE[(24*7):(24*7*2)], type="l")
```

Now we break up the timeseries into 7 day partitions, calculate the quantiles
at each time offset within the partitions to compute the range of weekly
patterns.

```{r hourlyquantilesplot}
weeks <- rollapply(all$TOTAL_COMPUTE, width=24*7, by=24*7,
                   FUN=function(x) return(x)) 

# Now we have a matrix.  Each row is one of the 1 week periods of time.
# Each 1-week period should be normalized separately to the peak demand that week.

weeks.norm <- apply(weeks, 1, function(x) { return ((100*x)/max(x)) })

# Now for each week, return a data frame and concat them together.
tmp.df <- do.call(rbind, lapply(1:ncol(weeks.norm), function(i) return (
  data.frame(x=weeks.norm[,i], y=1:length(weeks.norm[,i]),
             dow=rep(c("Su", "M", "T", "W", "R", "F", "S"), each=24, times=1)))))

# Now we generate a single data.frame of hourly quantiles.  One row for
# each hour in the week (24*7 = 168 rows) with a column for each of 7 different
# percentiles.

hourly.quantiles <- tmp.df %>% group_by(y) %>%
  summarize(p90 = quantile(x, probs=0.9),
            p95 = quantile(x, probs=0.95),
            p99 = quantile(x, probs=0.99),
            p10=quantile(x, probs=0.1),
            p5=quantile(x, probs=0.05),
            p1=quantile(x, probs=0.01),
            med=quantile(x, probs=0.5))

# Nice x-axis labels for the days of the week.
days <- c("Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat")

p3 <- ggplot(hourly.quantiles, aes(x=y, y=med)) +
  geom_ribbon(aes(ymin=p5, ymax=p95), fill="#DDE3EE") +
  geom_line(color="#5471AB") + theme_bw() +
  theme(axis.text = element_text(size=14),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16)) +
  xlab("Day of the Week") + ylab("Normalized Weekly\nVM Instance Demand") +
  scale_x_continuous(breaks=seq(from=0, to=24*14, length.out=14),
                     labels=rep(days, 2))

p3
```


```{r stackedpatternplot}
plot_grid(p2, p3, labels=c("(A)", "(B)"), scale=0.9, ncol=1)
```

```{r pdfout, eval=FALSE, fig.width=7, fig.height=5.5}
pdf("dailypattern.pdf", width=7, height=5.5)
plot_grid(p2, p3, labels=c("(A)", "(B)"), scale=0.9, ncol=1)
dev.off()
```

```{r pngout, eval=FALSE, fig.width=7, fig.height=5.5}
png("dailypattern.png", width=700, height=550)
plot_grid(p2, p3, labels=c("(A)", "(B)"), scale=0.9, ncol=1)
dev.off()
```

# Autocorrelation Analysis

```{r}
plot(all.meandaily)
pacf(all.meandaily)
#data.zoo <- zoo(all$NORM_USAGE, all$USAGE_HOUR)
#data.sub
#data.sub.zoo <- zoo(data.sub$TOTAL_COST, data.sub$day)
#data.sub.zoo
#plot(data.zoo)
#pacf(data.sub.zoo)
#acf(data.sub.zoo)
z = acf(all.meandaily)
z
y = acf(diff(all.meandaily))
y
x = acf(diff(all.meandaily,7))
x
```

# Hourly daily maximum vs daily minimum

```{r}
all.maxweekly <- rollapply(all.meandaily, width=7, by=7, FUN=max)
all.minweekly <- rollapply(all.meandaily, width=7, by=7, FUN=min)
head(all.maxweekly)
head(all.minweekly)
head(all.maxweekly / all.minweekly)
mean(all.maxweekly / all.minweekly)
```

```{r}
head(all.maxdaily)
head(all.mindaily)
head(all.maxdaily/all.mindaily)
mean(all.maxdaily/all.mindaily)
```
Average daily maximum is 34% higher than average daily minimum.

# Holiday Effect Analysis

```{r}
holidays.2022 <- subset(data.all, (USAGE_HOUR > as.POSIXct("2021-12-15") &
                                     USAGE_HOUR < as.POSIXct("2022-01-15")))
# Normalize this holiday period to 100 over the holidays.
holidays.2022$NORM_USAGE <- (100 * holidays.2022$NORM_USAGE) /
  max(holidays.2022$NORM_USAGE)
label.1 <- "2021-2022"
label.2 <- "2022-2023"
label.3 <- "2023-2024"
holidays.2022$year = label.1
holidays.2022$weekend.start = as.POSIXct("2021-12-18")
holidays.2022$weekend.end = as.POSIXct("2021-12-20")
holidays.2022$ts <- holidays.2022$USAGE_HOUR

highlights <- data.frame(year=c(rep(label.1, 4), rep(label.2, 5),
                                rep(label.3, 5)),
                         weekend.start=c(as.POSIXct("2021-12-18"),
                                         as.POSIXct("2021-12-25"),
                                         as.POSIXct("2022-01-01"),
                                         as.POSIXct("2022-01-08"),
                                         as.POSIXct("2021-12-17"),
                                         as.POSIXct("2021-12-24"),
                                         as.POSIXct("2021-12-31"),
                                         as.POSIXct("2022-01-07"),
                                         as.POSIXct("2022-01-14"),
                                         as.POSIXct("2021-12-16"),
                                         as.POSIXct("2021-12-23"),
                                         as.POSIXct("2021-12-30"),
                                         as.POSIXct("2022-01-06"),
                                         as.POSIXct("2022-01-13")),
                         weekend.end=c(as.POSIXct("2021-12-20"),
                                       as.POSIXct("2021-12-27"),
                                       as.POSIXct("2022-01-03"),
                                       as.POSIXct("2022-01-10"),
                                       as.POSIXct("2021-12-19"),
                                       as.POSIXct("2021-12-26"),
                                       as.POSIXct("2022-01-02"),
                                       as.POSIXct("2022-01-09"),
                                       as.POSIXct("2022-01-15"),
                                       as.POSIXct("2021-12-18"),
                                       as.POSIXct("2021-12-25"),
                                       as.POSIXct("2022-01-01"),
                                       as.POSIXct("2022-01-08"),
                                       as.POSIXct("2022-01-15")))

holidays.2023 <- subset(data.all, (USAGE_HOUR > as.POSIXct("2022-12-15") &
                                     USAGE_HOUR < as.POSIXct("2023-01-15")))
holidays.2023$ts <- holidays.2023$USAGE_HOUR - dyears(1)

holidays.2023$NORM_USAGE <- (100 * holidays.2023$NORM_USAGE) /
  max(holidays.2023$NORM_USAGE)
holidays.2023$year = label.2
holidays.2023$weekend.start = as.POSIXct("2021-12-17")
holidays.2023$weekend.end = as.POSIXct("2021-12-19")

holidays.2024 <- subset(data.all, (USAGE_HOUR > as.POSIXct("2023-12-15") &
                                     USAGE_HOUR < as.POSIXct("2024-01-15")))
holidays.2024$ts <- holidays.2024$USAGE_HOUR - dyears(2)

holidays.2024$NORM_USAGE <- (100 * holidays.2024$NORM_USAGE) /
  max(holidays.2024$NORM_USAGE)
holidays.2024$year = label.3
holidays.2024$weekend.start = as.POSIXct("2021-12-16")
holidays.2024$weekend.end = as.POSIXct("2021-12-18")

holidays.all <- rbind(holidays.2022, holidays.2023, holidays.2024)
head(holidays.all)
```



```{r}
holidayPlot <- function() {
  highlights$ts <- min(holidays.all$ts)
  highlights$NORM_USAGE <- min(holidays.all$NORM_USAGE)
  ggplot(data=holidays.all, aes(x=ts, y=NORM_USAGE))  +
    geom_rect(data=highlights, aes(xmin=weekend.start, xmax=weekend.end,
                                   ymin=-Inf, ymax=Inf),
              fill="#DDE3EE",alpha=1.0) +
    geom_line() + facet_grid(rows = vars(year)) + theme_bw() +
    theme(axis.text=element_text(size=15), axis.title=element_text(size=15),
          strip.text=element_text(size=15), legend.text = element_text(size=15),
          legend.title = element_text(size=15)) +
    ylab("Normalized VM Instance Demand") + xlab("") +
    scale_x_continuous(breaks=c(as.POSIXct("2021-12-18"), as.POSIXct("2021-12-25"),
                                as.POSIXct("2022-01-01"), as.POSIXct("2022-01-08"),
                                as.POSIXct("2022-01-15")),
                       labels=c("Dec 18", "Dec 25", "Jan 1", "Jan 8", "Jan 15"),
                       minor_breaks=seq(from=as.POSIXct("2021-12-19"),
                                        to=as.POSIXct("2022-01-15"), by=86400))
}
holidayPlot()
```

```{r holidaypdfout, echo=FALSE}
pdf("annualholiday.pdf", width=9, height=6)
holidayPlot()
dev.off()
```

```{r holidaypngout, echo=FALSE}
png("annualholiday.png", width=700, height=600*7/9)
holidayPlot()
dev.off()
```

