---
# Copyright 2025 Snowflake Inc. 
# SPDX-License-Identifier: Apache-2.0
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.
title: "Figures: Trend Analysis"
author: "Murray Stokely"
email: "murray.stokely@snowflake.com"
output:
  pdf_document:
    toc: true
    number_sections: true
  html_document:
    toc: true
    number_sections: true
date: "2025-01-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, results=FALSE, warning=FALSE, message=FALSE}
defaultW <- getOption("warn") # suppress all informational messages during pkg load
options(warn=-1)
library(zoo)
library(ggplot2)
library(tidyr)
library(dplyr)
library(arrow)
library(cowplot)
library(scales)
library(ggpattern) # for thatched area plots to not rely on color printing
options(warn=defaultW)
```

# Introduction

This file includes figures and analysis for Section 3.3.1 Trend in Setting Optimal
Commitment Level for Periodic Demand.

We start by aggregating the different regions and SKUs together to a single timeseries of total VM demand, and normalize this to a 100 unit peak over that time window.

```{r}
data <- read_parquet("../hourly_normalized.parquet")
dim(data)
head(data)
range(data$USAGE_HOUR)
data.all <- data %>% group_by(USAGE_HOUR) %>% summarise(across(NORM_USAGE, sum))
data.all$NORM_USAGE = 100*data.all$NORM_USAGE / max(data.all$NORM_USAGE)
head(data.all)
ggplot(data.all, aes(x=USAGE_HOUR, y=NORM_USAGE)) + geom_line()
```



# Weekly Pattern Analysis

We trim the start and end of our timeseries to align our dataset with Sunday through Saturday weeks so we can split it up in 7 day chunks and look at the distribution of weekly patterns.

We then generate timeseries of the maximum, minimum, and mean of each day.

```{r}
all <- data.all %>% subset(USAGE_HOUR >= as.POSIXct("2021-02-07", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2024-01-28", tz="UTC"))
range(all$USAGE_HOUR)
all.ts <- zoo(all$NORM_USAGE, all$USAGE_HOUR)
# Compute the maximumes over each day.
all.maxdaily <- rollapply(all.ts, width=24, by=24, FUN=max)
all.meandaily <- rollapply(all.ts, width=24, by=24, FUN=mean)
all.mindaily <- rollapply(all.ts, width=24, by=24, FUN=min)
plot(all.ts)
plot(all.maxdaily)
all.meandaily <- 100*all.meandaily/(max(all.meandaily))
plot(all.meandaily)
all.meanweekly <- rollapply(all.meandaily, width=7, by=7, FUN=mean)
plot(all.meanweekly)
```


```{r}
growthmultiple.3yr <- max(range(all.meanweekly)) / min(range(all.meanweekly))
growthrate.3yr <- growthmultiple.3yr ^ (1/3)
print(growthmultiple.3yr)
print(growthrate.3yr)
```

## Figure 5

Week over week growth timeseries.

```{r}
library(scales)
growth.rate <- diff(all.meanweekly) / head(all.meanweekly, -1)
growth.rate.df <- data.frame(time=time(growth.rate), value=as.numeric(growth.rate))
ggplot(growth.rate.df, aes(x=time, y=value)) + geom_bar(stat="identity", color="#5471AB") + theme_bw() + theme(axis.title.x=element_text(size=16), axis.title.y=element_text(size=16), axis.text=element_text(size=14)) + ylab("Week Over Week Growth") + xlab("") + scale_y_continuous(labels=scales::label_percent())
```

```{r echo=FALSE, eval=FALSE}
pdf("wowgrowth.pdf", width=7, height=3.25)
ggplot(growth.rate.df, aes(x=time, y=value)) + geom_bar(stat="identity", color="#5471AB") + theme_bw() + theme(axis.title.x=element_text(size=16), axis.title.y=element_text(size=16), axis.text=element_text(size=14)) + ylab("Week Over Week Growth") + xlab("") + scale_y_continuous(labels=scales::label_percent())
dev.off()
```


```{r echo=FALSE, eval=FALSE}
png("wowgrowth.png", width=700, height=325)
ggplot(growth.rate.df, aes(x=time, y=value)) + geom_bar(stat="identity", color="#5471AB") + theme_bw() + theme(axis.title.x=element_text(size=16), axis.title.y=element_text(size=16), axis.text=element_text(size=14)) + ylab("Week Over Week Growth") + xlab("") + scale_y_continuous(labels=scales::label_percent())
dev.off()
```

How many weeks is the trend negative?

```{r}
length(which(diff(all.meanweekly) < 0 )) / length(diff(all.meanweekly))
```

# Sensitivity Analysis of Trend and Period for Commitment Savings

First lets pick a week of data to start with:

```{r}
data.subset <- data.all %>%
  subset(USAGE_HOUR >= as.POSIXct("2024-01-21", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2024-01-28", tz="UTC"))
range(data.subset$USAGE_HOUR)
dim(data.subset)
```

Needed helper functions from optimization.Rmd

```{r echo=FALSE}
AnnotateSPRILevel <- function(df, sp.level) {
  df$SPRI_LEVEL = sp.level
  return(df)
}

GenerateBoxes <- function(df, on.demand.premium=2.1) {
  # GenerateBoxes - Create data.frame with pricing
  # Args:
  #   df: A data.frame with columns
  #     USAGE_HOUR: POSIXct hourly time
  #     NORM_USAGE: The VM demand at that hour
  # 3 boxes each time range - ununused, used with sp/ri coverage,
  #   used above sp/ri coverage level.
  data.boxes <- data.frame(
    xmin = rep(head(df$USAGE_HOUR, -1), 3),
    xmax = rep(tail(df$USAGE_HOUR, -1), 3),
    Pricing = c(rep("SP Commitment", length(head(df$USAGE_HOUR, -1))),
                rep("Unused Commitment", length(head(df$USAGE_HOUR, -1))),
                rep("On Demand", length(head(df$USAGE_HOUR, -1)))),
    ymin = c(rep(0, length(head(df$USAGE_HOUR, -1))),
             head(ifelse(df$NORM_USAGE < df$SPRI_LEVEL, df$NORM_USAGE,
                         df$SPRI_LEVEL), -1),
             head(df$SPRI_LEVEL, -1)),
    ymax = c(head(ifelse(df$NORM_USAGE < df$SPRI_LEVEL, df$NORM_USAGE,
                         df$SPRI_LEVEL), -1),
             head(df$SPRI_LEVEL,-1),
             head(ifelse(df$NORM_USAGE > df$SPRI_LEVEL,
                         df$SPRI_LEVEL +
                           (df$NORM_USAGE - df$SPRI_LEVEL)*on.demand.premium,
                         df$SPRI_LEVEL), -1)))
  
  return(data.boxes)
}

findMinSPLevel <- function(df, steps=9) {
  min.demand <- min(df$NORM_USAGE)
  max.demand <- max(df$NORM_USAGE)

  lowest.cost <- NULL
  lowest.sp <- NULL
  i<-1
  for (sp.level in seq(min.demand, max.demand, length=steps)) {
    df.ann <- AnnotateSPRILevel(df, sp.level)
    df.boxes <- GenerateBoxes(df.ann)
    df.boxes$sp.level.label <- sp.level
    cost.premium  <- TotalCostPremium(df.boxes)
    if (is.null(lowest.cost)) {
      lowest.cost <- cost.premium
      lowest.sp <- sp.level
    } else if (cost.premium < lowest.cost) {
      lowest.cost <- cost.premium
      lowest.sp <- sp.level
    }
  }
  return(lowest.sp)
}

findMinSPLevel.2 <- function(df, steps=NULL) {
  min.level <- min(df$NORM_USAGE)
  max.level <- max(df$NORM_USAGE)
  helper <- function(sp.level) {
    df.ann <- AnnotateSPRILevel(df, sp.level)
    df.boxes <- GenerateBoxes(df.ann)
    df.boxes$sp.level.label <- sp.level
    return(TotalCostPremium(df.boxes))
  }
  m <- optimize(helper, interval=c(min.level, max.level), tol=1e-20)
  return(m$minimum)
}


findMinSPLevel.3 <- function(df, steps=NULL, initial=NULL) {
  min.level <- min(df$NORM_USAGE)
  max.level <- max(df$NORM_USAGE)
  helper <- function(sp.level) {
    df.ann <- AnnotateSPRILevel(df, sp.level)
    df.boxes <- GenerateBoxes(df.ann)
    df.boxes$sp.level.label <- sp.level
    return(TotalCostPremium(df.boxes))
  }
  if (is.null(initial)) {
    initial = (min.level + max.level)/2
  }
  m <- optim(par=initial,
             fn=helper,
             lower=min.level,
             upper=max.level,
             control=list(maxit=5000))
  return(m$par)
}

TotalCostPremium <- function(df) {
  # TotalCost - Computes the Total Cost Premium
  #
  # Args:
  #   df - A data.frame with 3 areas - unused, covered, on-demand.
  # Returns
  #   Cost premium

  l1 <- subset(df, Pricing == "SP Commitment")
  l2 <- subset(df, Pricing == "Unused Commitment")
  l3 <- subset(df, Pricing == "On Demand")
  plan.cost <- sum(l1$ymax - l1$ymin)
  unused.cost <- sum(l2$ymax - l2$ymin)
  ondemand.cost <- sum(l3$ymax - l3$ymin)
  total.cost <- plan.cost + unused.cost + ondemand.cost
  return(total.cost)
}

FindCost <- function(df, sp.level) {
  df.ann <- AnnotateSPRILevel(df, sp.level)
  df.boxes <- GenerateBoxes(df.ann)
  df.boxes$sp.level.label <- sp.level
  return(TotalCostPremium(df.boxes))
}
```

```{r}
data.subset.orig <- data.subset
```

## Table 3

We now do a sensitivity analysis on the costs for different annual trend and number of weeks.

```{r}
data.subset <- data.subset.orig
#data.subset$NORM_USAGE <- data.subset$NORM_USAGE * 10

ExtendWeek <- function(df, weeks=1, trend=0.05) {
  daily.trend <- (1+trend)^(1/365)
  df.tmp <- df
  for (w in seq(weeks)) {
    df.new <- df
    df.new$USAGE_HOUR <- df.new$USAGE_HOUR + (w*7*86400)
    df.new$NORM_USAGE <- df.new$NORM_USAGE * daily.trend^(rep(seq(((w-1)*7)+1,((w-1)*7)+7),each=24))
    #df.new$DEMAND <- df.new$DEMAND * daily.trend^(rep(seq(((w-1)*7)+1,((w-1)*7)+7),each=24))
    df.tmp <- rbind(df.tmp, df.new) %>% arrange(USAGE_HOUR)
  }
  return(df.tmp)
}

# 100 steps was not enough to generate full table accurately.
steps <- 30000

for (annual.trend in c(.1, .25, .5, .75, 1)) {
  # Extend our dataset n weeks into future at specified annual trend
  df.week.1 <- ExtendWeek(data.subset, weeks=1, trend=annual.trend)
  df.week.2 <- ExtendWeek(data.subset, weeks=2, trend=annual.trend)
  df.week.4 <- ExtendWeek(data.subset, weeks=4, trend=annual.trend)
  df.week.8 <- ExtendWeek(data.subset, weeks=8, trend=annual.trend)

  # Filter out the actuals and only consider the future.
  df.forecast.1 <- df.week.1 %>% subset(USAGE_HOUR > max(data.subset$USAGE_HOUR))
  df.forecast.2 <- df.week.2 %>% subset(USAGE_HOUR > max(data.subset$USAGE_HOUR))
  df.forecast.4 <- df.week.4 %>% subset(USAGE_HOUR > max(data.subset$USAGE_HOUR))
  df.forecast.8 <- df.week.8 %>% subset(USAGE_HOUR > max(data.subset$USAGE_HOUR))

  # Find lowest cost SP on the actuals
  lowest.actuals <- findMinSPLevel.2(data.subset, steps)
  
  # Compute lowest cost on the forecast
  lowest.forecast.1 <- findMinSPLevel.2(df.forecast.1, steps)
  lowest.forecast.2 <- findMinSPLevel.2(df.forecast.2, steps)
  lowest.forecast.4 <- findMinSPLevel.2(df.forecast.4, steps)
  lowest.forecast.8 <- findMinSPLevel.2(df.forecast.8, steps)
  
  # Compute cost of the week1 forecast with our min SP level found from actuals
  cost.base.1 <- FindCost(df.forecast.1, lowest.actuals)
  cost.base.2 <- FindCost(df.forecast.2, lowest.actuals)
  cost.base.4 <- FindCost(df.forecast.4, lowest.actuals)
  cost.base.8 <- FindCost(df.forecast.8, lowest.actuals)
  cost.forecast.1 <- FindCost(df.forecast.1, lowest.forecast.1)
  cost.forecast.2 <- FindCost(df.forecast.2, lowest.forecast.2)
  cost.forecast.4 <- FindCost(df.forecast.4, lowest.forecast.4)
  cost.forecast.8 <- FindCost(df.forecast.8, lowest.forecast.8)
  
  print(paste0(" Annual Trend: ", annual.trend,
               sprintf(" CostBase:1 %0.7f", cost.base.1), sprintf(" Cost1Wk: %0.7f", cost.forecast.1),
               sprintf(" CostBase:2 %0.7f", cost.base.2), sprintf(" Cost2Wk: %0.7f", cost.forecast.2),
               sprintf(" CostBase:4 %0.7f", cost.base.4), sprintf(" Cost4Wk: %0.7f", cost.forecast.4),
               sprintf(" CostBase:8 %0.7f", cost.base.8), sprintf(" Cost8Wk: %0.7f", cost.forecast.8)))
  print(paste0(sprintf(" CostDelta.1: %0.7f", cost.base.1 - cost.forecast.1),
               sprintf(" CostDelta.2: %0.7f", cost.base.2 - cost.forecast.2),
               sprintf(" CostDelta.4: %0.7f", cost.base.4 - cost.forecast.4),
               sprintf(" CostDelta.8: %0.7f", cost.base.8 - cost.forecast.8)))
               
  print("")
}

for (weeks in c(1, 2, 4,8)) { #} c(1,2,4,8)) {
  for (annual.trend in c(.1, .25, .5, .75, 1)) {
    # Extend our dataset n weeks into future at specified annual trend
    df.tmp <- ExtendWeek(data.subset, weeks=weeks, trend=annual.trend)
    # Find lowest cost on the actuals
    lowest.actuals <- findMinSPLevel.2(data.subset, steps)
    # Filter out the actuals and only consider the future.
    df.forecast <- df.tmp %>% subset(USAGE_HOUR > max(data.subset$USAGE_HOUR))
    # Compute lowest cost on the forecast
    lowest.forecast <- findMinSPLevel.2(df.forecast, steps)
    # Compute cost of the forecast with our min SP level found from actuals
    cost.base <- FindCost(df.forecast, lowest.actuals)
    # Compute cost of the forecast with our min SP level found from forecast
    cost.forecast <- FindCost(df.forecast, lowest.forecast)
    #
    extra.cost <- ((1000000*cost.base) / cost.forecast) - 1000000
    #extra.cost <- ((cost.1 / cost.forecast) * 1000000) - 1000000
    print(paste0(" Weeks: ", weeks, sprintf(" MaxVal: %0.1f", max(df.tmp$NORM_USAGE)), sprintf(" Trend: %0.2f", annual.trend),
                 sprintf(" LowActuals: %0.3f", lowest.actuals),
                 sprintf(" LowForecast: %0.7f", lowest.forecast),
                 sprintf(" BaseCost: %0.3f", cost.base),
                 sprintf(" Cost2: %0.3f", cost.forecast),
                 sprintf(" Extra Cost: %0.4f", extra.cost),
                 sprintf(" CostDelta: %0.7f", (cost.base - cost.forecast)),
                 #sprintf(" CostDeltaPct: %0.4f", 100*(cost.base - cost.forecast)/cost.base),
                 sprintf(" CostPM: %0.4f", (1000000/cost.base) * (cost.base-cost.forecast))))
  }
  print("")
}

df.tmp <- ExtendWeek(data.subset, weeks=1, trend=0.05)
lowest.1 <- findMinSPLevel(data.subset, 100)
df.forecast <- df.tmp %>% subset(USAGE_HOUR > max(data.subset$USAGE_HOUR))
lowest.forecast <- findMinSPLevel(df.forecast, 100)
cost.1 <- FindCost(df.forecast, lowest.1)
cost.forecast <- FindCost(df.forecast, lowest.forecast)
cost.1
cost.forecast
```


### Pull in PlotBoxes from optimization.Rmd

```{r}

PlotBoxes <- function(df, sp.level=NULL, title=FALSE, ylim=NULL) {
  days <- c("S", "M", "T", "W", "R", "F", "S")
  p <- ggplot(df) +
           geom_rect_pattern(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax,
                                 fill=Pricing, pattern=Pricing),
                             colour=NA, pattern_size=0.25,
                             pattern_spacing=0.02) +
           theme_bw() +
           theme(axis.text=element_text(size=15),
                 axis.title=element_text(size=15),
                 legend.text = element_text(size=15),
                 legend.title = element_blank(),
                 legend.position="bottom",
                 strip.text=element_text(size=15)) +
           ylab("Cost") + xlab("") +
           scale_pattern_manual(values=c("none", "none", "stripe")) +
           scale_x_continuous(breaks=seq(from=as.POSIXct("2024-06-02 12:00:00"),
                                         to=as.POSIXct("2024-06-15 12:00:00"),
                                         length.out=14),
                              labels=rep(days,2)) +
           scale_fill_manual(values=c("orange", "darkgreen", "red"))
  if (title) {
    if (!is.null(sp.level)) {
      cost.premium  <- TotalCostPremium(df)
      p <- p + ggtitle(paste0("c=", round(sp.level, 1), " ", "C(c)=", round(cost.premium, 0)))
    }
  }
  if (!is.null(ylim)) {
    p <- p+ylim(ylim)
  }
  return(p)
}
```

## Figure 6

We create a new data frame from our existing week extended out 8 weeks with 100% annual trend and compare the
SP level that would be set on the initial data vs the 8-week forecast.

```{r}
df.tmp<-ExtendWeek(data.subset, weeks=8, trend=1)
lowest.1 <- findMinSPLevel(data.subset, 100)
df.forecast <- df.tmp %>% subset(USAGE_HOUR > max(data.subset$USAGE_HOUR))
lowest.forecast <- findMinSPLevel(df.forecast, 100)
cost.1 <- FindCost(df.forecast, lowest.1)
cost.forecast <- FindCost(df.forecast, lowest.forecast)
cost.1
cost.forecast

PlotBest.2 <- function(df, sp.level, annotated.level) {
  df.ann <- AnnotateSPRILevel(df, sp.level)
  df.boxes <- GenerateBoxes(df.ann)
  brks <- seq(from=min(df.boxes$xmin), to=max(df.boxes$xmax), by=(60*60*24*7))
  p <- PlotBoxes(df.boxes)
  p <- p +
    scale_x_continuous(breaks=brks, labels=1:length(brks)) +
    geom_hline(aes(yintercept=sp.level,
                   linetype="Min Cost Commitment for 1-week"),
               color="black") +
    geom_hline(aes(yintercept=annotated.level,
                   linetype="Min Cost Commitment for 8-weeks"),
               color="red") +
    scale_linetype_manual(name="Commitment Level",
                          values=c("dashed", "dashed"),
                          guide=guide_legend(
                            position="top",
                            direction="vertical",
                            override.aes = list(color=c("black", "red")))) +
    xlab("Week")
  return(p)
}

p <- PlotBest.2(df.forecast, lowest.1, lowest.forecast)
p
```

```{r echo=FALSE,eval=FALSE}
pdf("trendforecast.pdf", width=7, height=4)
p
dev.off()
```

```{r echo=FALSE,eval=FALSE}
png("trendforecast.png", width=700, height=400)
p
dev.off()
```



# Laddering

We consider the 4 weeks between 12/3 and 12/31/2023 to consider the impact of laddering.

```{r}
PlotBoxes.2 <- function(df, sp.level=NULL, title=FALSE, ylim=NULL, weeks=1) {
  daysdelta = round(as.numeric(max(df$xmax) - min(df$xmin)))
  p <- ggplot(df) +
           geom_rect_pattern(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax,
                                 fill=Pricing, pattern=Pricing),
                             colour=NA, pattern_size=0.25,
                             pattern_spacing=0.02) +
           theme_bw() +
           theme(axis.text=element_text(size=15),
                 axis.title=element_text(size=15),
                 legend.text = element_text(size=15),
                 legend.title = element_blank(),
                 legend.position="bottom",
                 strip.text=element_text(size=15)) +
           ylab("Cost") + xlab("") +
           scale_pattern_manual(values=c("none", "none", "stripe")) +
           scale_x_continuous(breaks=seq(from=as.POSIXct(min(df$xmin), tz="UTC"),
                                         to=as.POSIXct(max(df$xmax), tz="UTC"),
                                         length.out=5),
                              labels=format(as.Date(seq(from=as.POSIXct(min(df$xmin), tz="UTC"),
                                         to=as.POSIXct(max(df$xmax), tz="UTC"),
                                         length.out=5)), "%b-%d")) +
#                                         to=min(df$xmin) + daysdelta*86400 - 60*60*12,
#                                         length.out=daysdelta),
#                              labels=rep(days, round(daysdelta/7))) +
           scale_fill_manual(values=c("orange", "darkgreen", "red"))
  if (title) {
    if (!is.null(sp.level)) {
      cost.premium  <- TotalCostPremium(df)
      p <- p + ggtitle(paste0("c=", round(sp.level, 1), " ", "C(c)=", round(cost.premium, 0)))
    }
  }
  if (!is.null(ylim)) {
    p <- p+ylim(ylim)
  }
  return(p)
}
```

```{r}
data.tmp <- data.all %>% 
  subset(USAGE_HOUR >= as.POSIXct("2023-12-03", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2023-12-31", tz="UTC"))
# Re-normalize just these 4 weeks to 100
data.tmp$NORM_USAGE <- data.tmp$NORM_USAGE * (100/max(data.tmp$NORM_USAGE))
data.w1 <- subset(data.tmp, USAGE_HOUR < as.POSIXct("2023-12-10", tz="UTC"))
data.w2 <- subset(subset(data.tmp, USAGE_HOUR < as.POSIXct("2023-12-17", tz="UTC")),
                  USAGE_HOUR >= as.POSIXct("2023-12-10", tz="UTC"))
data.w3 <- subset(subset(data.tmp, USAGE_HOUR < as.POSIXct("2023-12-24", tz="UTC")),
                  USAGE_HOUR >= as.POSIXct("2023-12-17", tz="UTC"))
data.w4 <- subset(subset(data.tmp, USAGE_HOUR < as.POSIXct("2024-01-01", tz="UTC")),
                  USAGE_HOUR >= as.POSIXct("2023-12-24", tz="UTC"))
minlev.1 <- findMinSPLevel(data.w1, steps=100)
minlev.2 <- findMinSPLevel(data.w2, steps=100)
minlev.3 <- findMinSPLevel(data.w3, steps=100)
minlev.4 <- findMinSPLevel(data.w4, steps=100)
minlev <- findMinSPLevel(data.tmp, steps=100)
data.1 <- AnnotateSPRILevel(data.tmp, sp.level=minlev.4)
data.1$SPRI_LEVEL[data.1$USAGE_HOUR < as.POSIXct("2023-12-24", tz="UTC")] <- minlev.3
data.1$SPRI_LEVEL[data.1$USAGE_HOUR < as.POSIXct("2023-12-17", tz="UTC")] <- minlev.2
data.1$SPRI_LEVEL[data.1$USAGE_HOUR < as.POSIXct("2023-12-10", tz="UTC")] <- minlev.1
data.b <- GenerateBoxes(data.1)
show(PlotBoxes.2(data.b))
print(minlev.1)
```

```{r}
data.one <- AnnotateSPRILevel(data.tmp, sp.level=minlev)
data.oneb <- GenerateBoxes(data.one)
print(TotalCostPremium(data.oneb))
show(PlotBoxes.2(data.oneb))

```

```{r}
print(TotalCostPremium(data.b))
minlev.1
```


```{r}
TotalCostPremium(data.b) / TotalCostPremium(data.oneb)
```


```{r plot3by3}
Plot3x3 <- function(df, start.date="2024-01-07", end.date="2024-01-21", weekdays=FALSE) {
  days <- c("S", "M", "T", "W", "R", "F", "S")
  p <- ggplot(df) +
    geom_rect_pattern(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax,
                          fill=Pricing, pattern=Pricing),
                      colour=NA, pattern_size=0.25, pattern_spacing=0.02) +
    theme_bw() +
    theme(axis.text=element_text(size=15), axis.title=element_text(size=15),
     #     axis.text.x = element_text(angle=90),
          legend.text = element_text(size=15), legend.title = element_blank(),
          legend.position="bottom", strip.text=element_text(size=15)) +
    ylab("Cost") + xlab("") +
    scale_pattern_manual(values=c("none", "none", "stripe"))
  if (weekdays) {
    p <- p + scale_x_continuous(breaks=seq(from=as.POSIXct(start.date, tz="UTC"),
                                  to=as.POSIXct(end.date, tz="UTC"),
                                  length.out=14),
                      labels=rep(days,2))
  } else {
     p <- p + scale_x_continuous(breaks=c(as.POSIXct("2023-12-03", tz="UTC"),
                                          as.POSIXct("2023-12-10", tz="UTC"),
                                          as.POSIXct("2023-12-17", tz="UTC"),
                                          as.POSIXct("2023-12-24", tz="UTC")),
                                 labels=date_format("%b-%d"))
#                                          seq(from=as.POSIXct(min(df$xmin), tz="UTC"),
#                                         to=as.POSIXct(max(df$xmax), tz="UTC"),
#                                         length.out=5),
#                              labels=format(as.Date(seq(from=as.POSIXct(min(df$xmin), tz="UTC"),
#                                         to=as.POSIXct(max(df$xmax), tz="UTC"),
                                         #length.out=5)), "%b-%d")) 
    #p <- p + scale_x_datetime(breaks = "1 week", minor_breaks = "1 day", labels=date_format("%b-%d"))
    #, limits=c(min(df$xmin), max(df$xmax)))
  }
  p <- p +
    scale_fill_manual(values=c("orange", "darkgreen", "red")) +
    facet_wrap(~panel.title)
  return(p)
}
```

## Figure 9

```{r}
data.oneb$panel.title <- paste0("(a) c=", round(minlev, 1), " C(c)=", round(TotalCostPremium(data.oneb), 0))
data.b$panel.title <- paste0("(b) c=(", round(minlev.1, 0), ", ", round(minlev.2, 0), ", ", round(minlev.3, 0), ", ", round(minlev.4, 0), ") C(c)=", round(TotalCostPremium(data.b), 0))

fig9 <- rbind(data.b, data.oneb)
Plot3x3(fig9, start.date="2023-12-17", end.date="2023-12-31")
```

```{r echo=FALSE,eval=FALSE}
pdf("ladderedholidayex.pdf", width=9, height=3.5)
Plot3x3(fig9, start.date="2023-12-17", end.date="2023-12-31")
dev.off()
```


```{r echo=FALSE,eval=FALSE}
png("ladderedholidayex.png", width=900, height=350)
Plot3x3(fig9, start.date="2023-12-17", end.date="2023-12-31")
dev.off()
```

