---
# Copyright 2025 Snowflake Inc. 
# SPDX-License-Identifier: Apache-2.0
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.
title: 'Animation: Identifying Minimal Cost Compute Commitment Level'
author: "Murray Stokely"
email: "murray.stokely@snowflake.com"
output:
  pdf_document:
    toc: true
    number_sections: true
  html_document:
    toc: true
    number_sections: true
date: "2024-10-08"
---


```{r echo=FALSE, results=FALSE, warning=FALSE, message=FALSE}
defaultW <- getOption("warn") # suppress all informational messages during pkg load
options(warn=-1)
library(zoo)
library(Cairo)
library(ggplot2)
library(tidyr)
library(dplyr)
library(arrow)
library(cowplot)
library(ggpattern) # for thatched area plots to not rely on color printing
options(warn=defaultW)
```


# Introduction

This file includes code to generate a series of hundreds or thousands of figures to serve as frames
in an animation of the optimization process.

We start by restricting to a 2-week subset of the data from January 2024, aggregate the different regions and VM types together to a single timeseries of total VM demand, and normalize this to a 100 unit peak over that time window.

```{r readdata}
this.dir <- getwd()
data <- read_parquet("../hourly_normalized.parquet")
data <- data %>% subset(USAGE_HOUR >= as.POSIXct("2024-01-07", tz="UTC")) %>%
  subset(USAGE_HOUR < as.POSIXct("2024-01-21", tz="UTC"))
dim(data)
head(data)
range(data$USAGE_HOUR)
data.all <- data %>% group_by(USAGE_HOUR) %>% summarise(across(NORM_USAGE, sum))
data.all$NORM_USAGE = 100 * data.all$NORM_USAGE / max(data.all$NORM_USAGE)
head(data.all)
ggplot(data.all, aes(x=USAGE_HOUR, y=NORM_USAGE)) + geom_line()
```

# Visualization and Cost Computations

In order to illustrate the impact of savings plans on the amount of VM demand
that is (1) covered by a savings plan, (2) purchased with on-demand rates, and (3) wasted as an unused savings plan,
we introduce a simple 3-color area visualization with time on the x-axis and normalized cost on the y-axis.

Note that since the y-axis is cost, instead of VM instance hours, the more of the demand that is covered at expensive on-demand rates means the higher the y-axis will be.

```{r helperfunctions}
TotalCostPremium <- function(df) {
  # TotalCost - Computes the Total Cost Premium
  #
  # Args:
  #   df - A data.frame with 3 areas - unused, covered, on-demand.
  # Returns
  #   Cost premium

  l1 <- subset(df, Pricing == "SP Commitment")
  l2 <- subset(df, Pricing == "Unused Commitment")
  l3 <- subset(df, Pricing == "On Demand")
  plan.cost <- sum(l1$ymax - l1$ymin)
  unused.cost <- sum(l2$ymax - l2$ymin)
  ondemand.cost <- sum(l3$ymax - l3$ymin)
  total.cost <- plan.cost + unused.cost + ondemand.cost
  return(total.cost)
}

GenerateBoxes <- function(df, on.demand.premium=2.1, label=NULL) {
  # GenerateBoxes - Create data.frame with pricing
  # Args:
  #   df: A data.frame with columns
  #     USAGE_HOUR: POSIXct hourly time
  #     NORM_USAGE: The VM demand at that hour
  # 3 boxes each time range - ununused, used with sp/ri coverage,
  # used above sp/ri coverage level.
  data.boxes <- data.frame(
    xmin = rep(head(df$USAGE_HOUR, -1), 3),
    xmax = rep(tail(df$USAGE_HOUR, -1), 3),
    Pricing = c(rep("SP Commitment", length(head(df$USAGE_HOUR, -1))),
                rep("Unused Commitment", length(head(df$USAGE_HOUR, -1))),
                rep("On Demand", length(head(df$USAGE_HOUR, -1)))),
    ymin = c(rep(0, length(head(df$USAGE_HOUR, -1))),
             head(ifelse(df$NORM_USAGE < df$SPRI_LEVEL, df$NORM_USAGE,
                         df$SPRI_LEVEL), -1), head(df$SPRI_LEVEL, -1)),
    ymax = c(head(ifelse(df$NORM_USAGE < df$SPRI_LEVEL,
                         df$NORM_USAGE, df$SPRI_LEVEL), -1),
             head(df$SPRI_LEVEL,-1),
             head(ifelse(df$NORM_USAGE > df$SPRI_LEVEL,
                         df$SPRI_LEVEL +
                           (df$NORM_USAGE - df$SPRI_LEVEL)*on.demand.premium,
                         df$SPRI_LEVEL), -1)))
  
  if (!is.null(label)) {
    data.boxes$LABEL = label
  }
  return(data.boxes)
}

AnnotateSPRILevel <- function(df, sp.level) {
  df$SPRI_LEVEL = sp.level
  return(df)
}

# DEMAND = sp + ri + (ondemand / on.demand.premium) - unused
# sp + ri + ondemand = 41567.9
# 
PlotBoxes <- function(df, label.size=15, label=NULL, ymax=NULL,
                      hide.legend=TRUE) {
  days <- c("S", "M", "T", "W", "R", "F", "S")
  p <- ggplot(df) +
    geom_rect_pattern(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax,
                          fill=Pricing, pattern=Pricing), colour=NA,
                      pattern_size=0.25, pattern_spacing=0.02) +
    theme_bw() +
    theme(axis.text=element_text(size=label.size),
          title=element_text(size=1.2*label.size),
          axis.title=element_text(size=label.size),
          legend.text = if (hide.legend) element_blank() else
            element_text(size=label.size),
          legend.title = if (hide.legend) element_blank() else
            element_text(size=label.size),
          legend.position="bottom", strip.text=element_text(size=label.size)) +
    ylab("Cost") + xlab("") +
    scale_pattern_manual(values=c("none", "none", "stripe")) +
    scale_x_continuous(breaks=seq(from=as.POSIXct("2024-06-02 12:00:00"),
                                  to=as.POSIXct("2024-06-15 12:00:00"),
                                  length.out=14),
                       labels=rep(days,2)) +
    scale_fill_manual(values=c("orange", "darkgreen", "red"))
  if (!is.null(ymax)) {
    p <- p + ylim(c(0,ymax))
  }
  
  if (is.null(label)) {
    return(p + ggtitle(paste0("Biweekly Cost: $",
                              round(TotalCostPremium(df)/1000, 2), "K")))
  } else {
    return(p + facet_wrap(~LABEL) + theme(strip.text = element_text(size=16)))
  }
}
```

  
```{r generatesingleplot}
df <- AnnotateSPRILevel(data.all,
                        (max(data.all$NORM_USAGE) + min(data.all$NORM_USAGE))/2)
data.boxes <- GenerateBoxes(df)

PlotBoxes(data.boxes)
```

# Comparing Commitment Level (x-axis) vs Total Cost (y-axis)

As we iterate through many possible solutions to identify the lowest cost option
we keep track and plot the different commitment levels and the associated total cost.

```{r optimizationcurve}
xs <- seq(min(data.all$NORM_USAGE), max(data.all$NORM_USAGE), length=10)
ys <- c()
for (commit.level in xs) {
  df.tmp <- AnnotateSPRILevel(data.all, sp.level=commit.level)
  data.boxes.tmp <- GenerateBoxes(df.tmp, label=commit.level)
  cost.tmp <- TotalCostPremium(data.boxes.tmp)
  ys <- c(ys, cost.tmp)
}

opt.curve <- data.frame(x=xs, y=ys)
ggplot(opt.curve, aes(x=x, y=y)) + geom_line() + theme_bw() +
  xlab("Savings Plan Commitment Level") + ylab("Total Biweekly Cost") +
  geom_point(x=xs[1], y=ys[1], size=3, color="red")
```

# Animations

## Commitment Level Area Plot

Our basic approach for generating animations is to create a number of plots in a loop with
identical x and y axis limits and then stitch them together with ffmpeg into an animation.

We first look at our area plot showing the impact of different commitment levels on the amount of
used commitment, unused commitment, and on demand capacity.

The individual frames are written out to an ephermal temporary directory created by the R session,
but the final output animations are written out to the current directory.

For a smoother animation, create more frames by increasing the steps number below.

```{r areaplotframes}
#steps <- 100
steps <- 10
ymin <- NULL
ymax <- NULL
ls <- round(seq(from=min(data.all$NORM_USAGE), to=max(data.all$NORM_USAGE), length=steps))
i <- 1
tdir <- tempdir()

for (commit.level in ls) {
  df.sub.tmp <- AnnotateSPRILevel(data.all, sp.level=commit.level)
  data.boxes.tmp <- GenerateBoxes(df.sub.tmp, label=commit.level)
  y <- TotalCostPremium(data.boxes.tmp)
  if (is.null(ymin)) {
    ymin <- y
  } else if (y < ymin) {
    ymin <- y
  }
  if (is.null(ymax)) {
    ymax <- y
  } else if (y > ymax) {
    ymax <- y
  }
  filename <- file.path(tdir, paste0(sprintf("%03d", i), ".png"))
  print(paste0("Working on ", commit.level, " filename: ", filename))
  i <- i + 1
  #CairoPNG(filename, width=640, height=480)
  CairoPNG(filename, width=3840, height=2160)
  show(PlotBoxes(data.boxes.tmp, label.size=100, ymax=150))
  dev.off()
}
```

```{r firstanimation}
input.files <- file.path(tdir, "%03d.png")
out.file <- file.path(this.dir, "out.mp4")
cmd <- paste0("ffmpeg -y -r 10 -i ", input.files,
              " -c:v libx264 -r 30 -pix_fmt yuv420p ", out.file)
system(cmd)
```

## Cost vs Commitment Level Optimization Curve

We then look at generating some static images for the cost vs commitment level curve.

```{r curveframes}
steps <- 10
label.size <- 10
xs <- seq(min(data.all$NORM_USAGE), max(data.all$NORM_USAGE), length=10)
i <- 1

tdir <- tempdir()
x <- c()
ys <- c()
xmin <- min(data.all$NORM_USAGE)
xmax <- max(data.all$NORM_USAGE)

for (commit.level in xs) {
  df.tmp <- AnnotateSPRILevel(data.all, sp.level=commit.level)
  data.boxes.tmp <- GenerateBoxes(df.tmp, label=commit.level)
  cost.tmp <- TotalCostPremium(data.boxes.tmp)
  ys <- c(ys, cost.tmp)
  x <- c(x, commit.level)
  # Now make a graph
  opt.curve <- data.frame(x=x, y=ys)
  
  
  filename <- file.path(tdir, paste0(sprintf("s%03d", i), ".png"))
  i <- i+1
  print(paste0("Working on ", filename))
  label.size <- 100
  CairoPNG(filename, width=3840, height=2160)

  show(ggplot(opt.curve, aes(x=x, y=y)) + geom_line() + theme_bw() +
         theme(axis.text=element_text(size=label.size),
               axis.title=element_text(size=label.size)) +
         ylim(c(.95*ymin, 1.05*ymax)) +
         xlim(c(xmin, xmax)) +
         xlab("Savings Plan Commitment Level") +
         ylab("Total Cost to Snowflake") +
         geom_point(x=commit.level, y=cost.tmp, size=label.size*.5,
                    color="red"))
  dev.off()
}
```

## Combined Animation with both plots

The final step is to combine both plots side by side in the animation to see the impact on total cost spend as the savings plan commitment level changes.

```{r combinedplotframes}
#steps <- 600
steps <- 10
min.commit <- min(data.all$NORM_USAGE)
max.commit <- max(data.all$NORM_USAGE)
commit.levels <- round(seq(from=min(data.all$NORM_USAGE),
                           to=max(data.all$NORM_USAGE), length=steps))
print(commit.levels)
i <- 1
curr.d <- getwd()

x <- c()
ys <- c()
label.size <- 100
min.cost <- NULL
min.level <- NULL

for (commit.level in commit.levels) {
  df.sub.tmp <- AnnotateSPRILevel(data.all, sp.level=commit.level)
  data.boxes.tmp <- GenerateBoxes(df.sub.tmp, label=commit.level)
  cost.tmp <- TotalCostPremium(data.boxes.tmp)
  if (is.null(min.cost)) {
    min.cost <- cost.tmp
    min.level <- commit.level
  }
  if (min.cost > cost.tmp) {
    min.cost <- cost.tmp
    min.level <- commit.level
  }
  ys <- c(ys, cost.tmp)
  x <- c(x, commit.level)

  filename <- file.path(tdir, paste0(sprintf("c%04d", i), ".png"))
  print(paste0("Working on ", commit.level, " filename: ", filename))
  i <- i + 1
  #CairoPNG(filename, width=640, height=480)
  CairoPNG(filename, width=3840, height=2160)
  p1 <- PlotBoxes(data.boxes.tmp, label.size=label.size*0.9, ymax=150)
  opt.curve <- data.frame(x=x, y=ys)
  p2 <- ggplot(opt.curve, aes(x=x, y=y)) + geom_line() + theme_bw() +
    theme(axis.text=element_text(size=label.size*0.8),
          axis.title=element_text(size=label.size*0.8)) +
    ylim(c(.98*ymin, 1.02*ymax)) + xlim(c(xmin, xmax)) +
    xlab("Hourly Commitment Level") + ylab("Total Biweekly Cost") +
    geom_point(x=commit.level, y=cost.tmp, size=label.size*.5, color="red") +
    theme(plot.margin=unit(c(500,50,300,400), "points"))
  
  show(plot_grid(p1, p2, labels = c('A', 'B'), label_size = 12))
  dev.off()
}
```

```{r combinedanimation}
input.files <- file.path(tdir, "c%04d.png")
out.file <- file.path(this.dir, "combined.mp4")
cmd <- paste0("ffmpeg -y -i ", input.files,
              " -c:v libx264 -r 24 -pix_fmt yuv420p ", out.file)
system(cmd)
```